---
title: "Assignment 3 - Group Project"
author: "Team G"
date: "11 December 2017"
output: html_document
---

####2.1 Pre-Processing

Refer to the Loans_processed.csv file to confirm these instructions have been implemented

####2.2 Prediction

##### Question 1
```{r}

rm(list=ls())
new_data <- read.csv('C:/Users/karim/Documents/Imperial/Machine Learning/ProblemSets/Assignment3/Loans_processed.csv')
head(new_data)
summary(new_data)

set.seed(123)
shuffle_data <- new_data[sample(nrow(new_data)),]
training <- shuffle_data[1:20000,]
validation <- shuffle_data[20001:28000,]
test <- shuffle_data[28001:nrow(shuffle_data),]

prop.table(table(training$loan_status))
table(validation$loan_status)


```

######Question 2

```{r}
library(C50)
library(gmodels)


#Calculating accuracy on the total sample
accuracytab <- table(new_data$loan_status)
accuracy_num <- as.vector(accuracytab)
accuracy <- accuracy_num[2]/(accuracy_num[1]+accuracy_num[2])
error <- 1-accuracy
print(accuracy)
print(error)

train_data <- C5.0(training[,-8],training[,8])
train_outcome <- predict(train_data, training, type = "class")
summary(train_data)
table(train_outcome)

validation_outcome <- predict(train_data, validation, type="class")
table(validation_outcome)

CrossTable(training$loan_status, train_outcome, prop.chisq = FALSE, prop.c = FALSE, prop.r = FALSE, dnn = c('actual default', 'predicted default'))

CrossTable(validation$loan_status, validation_outcome, prop.chisq = FALSE, prop.c = FALSE, prop.r = FALSE, dnn = c('actual default', 'predicted default'))



```
Assessing our total sample, we have an accuracy level of 85.9% with a corresponding misclassification error of 14.1%.

When running the C50 algorithm, this only predicts that all loans will be fully repaid. The corresponding errors shown within the confusion matrix, reflects how none of the charged-off loans have been correctly predicted.

######Question 3

```{r}

#To obtain results that generate a sensitivity of c.25% on charged-off loans
cost_matrix25 <- matrix(c(NA, 1,
                         3.15, NA), 2, byrow=TRUE)
train_data25 <- C5.0(training[,-8],training[,8], costs = cost_matrix25)

validation_outcome25 <- predict(train_data25, validation, type="class")
CrossTable(validation$loan_status, validation_outcome25, prop.chisq = FALSE, prop.c = FALSE, prop.r = FALSE, dnn = c('actual default', 'predicted default'))

output25 <- table(validation[,8],validation_outcome25, dnn=c("actual", "predicted"))
kable(output25$byClass, caption = "Important Ratios")


sensitivity25 <- 276 / (276+821)
print(sensitivity25)

#To obtain results that generate a sensitivity of c.40% on charged-off loans
cost_matrix40 <- matrix(c(NA, 1,
                         3.7, NA), 2, byrow=TRUE)
train_data40 <- C5.0(training[,-8],training[,8], costs = cost_matrix40)

validation_outcome40 <- predict(train_data40, validation, type="class")
x40 <- CrossTable(validation$loan_status, validation_outcome40, prop.chisq = FALSE, prop.c = FALSE, prop.r = FALSE, dnn = c('actual default', 'predicted default'))

x40$tbl['FULLY PAID', 'FULLY PAID']

sensitivity40 <- 437 / (437+660)
print(sensitivity40)


#To obtain results that generate a sensitivity of c.50% on charged-off loans
cost_matrix50 <- matrix(c(NA, 1,
                         4.7, NA), 2, byrow=TRUE)
train_data50 <- C5.0(training[,-8],training[,8], costs = cost_matrix50)

validation_outcome50 <- predict(train_data50, validation, type="class")
x40 <- CrossTable(validation$loan_status, validation_outcome50, prop.chisq = FALSE, prop.c = FALSE, prop.r = FALSE, dnn = c('actual default', 'predicted default'))


sensitivity50 <- 561 / (561+536)
print(sensitivity50)





```

#####Question 4

```{r}

cost_matrix_opt <- matrix(c(NA, 1,
                         2.25, NA), 2, byrow=TRUE)
train_data_opt <- C5.0(training[,-8],training[,8], costs = cost_matrix_opt)

validation_outcome_opt <- predict(train_data_opt, validation, type="class")
CrossTable(validation$loan_status, validation_outcome_opt, prop.chisq = FALSE, prop.c = FALSE, prop.r = FALSE, dnn = c('actual default', 'predicted default'))

misclasserror <- (999+266) / (999+266+98+6637)
print(misclasserror)

```

This depends heavily on our level of risk aversion. A higher multiple will help reduce false positives (predicted repayments against actual charge-off), at the cost of far higher false negatives (i.e. loans predicted to generate a charge-off compared to actual repayment). The latter will result in loans not being made, undermining potential profitability. 

It is assumed striking a balance between both will require us to minimise our misclassification error. Hence applying a multiple of 2x weighting to loans in default, can help minimise this, albeit warranting at least a minimum of 991 applications being reviewed.


######Question 5

```{r}

cost_matrix_A <- matrix(c(NA, 1,
                         4, NA), 2, byrow=TRUE)
train_data_A <- C5.0(training[,-8],training[,8], costs = cost_matrix_A)

test_outcome_A <- predict(train_data_A, test, type="class")
CrossTable(test$loan_status, test_outcome_A, prop.chisq = FALSE, prop.c = FALSE, prop.r = FALSE, dnn = c('actual default', 'predicted default'))


cost_matrix_opt <- matrix(c(NA, 1,
                         2.25, NA), 2, byrow=TRUE)
train_data_opt <- C5.0(training[,-8],training[,8], costs = cost_matrix_opt)

test_outcome_opt <- predict(train_data_opt, test, type="class")
CrossTable(test$loan_status, test_outcome_opt, prop.chisq = FALSE, prop.c = FALSE, prop.r = FALSE, dnn = c('actual default', 'predicted default'))


```

