---
title: "Machine Learning Group Assignment 3"
author: "Qiuyue Zhang 01378527"
date: "10/12/2017"
output:
  html_document: default
  pdf_document: default
---

```{r, include=FALSE}
library(C50)
library(knitr)
library(lattice)
library(ggplot2)
library(e1071)
library(caret)
library(dplyr)
library(reshape)
```


1. Import the pre-processed data set in R. Shuffle the records and split them into a training set (20,000 records), a validation set (8,000 records) and a test set (all remaining records).

```{r}
loan <- read.csv('/Users/Veronique/Desktop/ICBA/Nov_Dec Term/Machine Learning/Assignment/Assignment3/Loans_processed.csv')
set.seed(1)
loan_shuffled <- loan[sample(1:nrow(loan)), ]
training_set <- loan_shuffled[1:20000, ]
validation_set <- loan_shuffled[20001:28000, ]
test_set <- loan_shuffled[28001:nrow(loan), ]
label <- c("Charged Off", "Fully Paid")
```


2. Using a classification tree (look at the C50 library), try to predict with an accuracy greater than $\frac {# of repaid loans}{# of repaid loans + # of charged off loans}$ if a loan will be repaid. Do you manage to achieve this performance on the validation set? What about the training set?

```{r}
# training
loan_tree_model <- C50::C5.0(x = training_set[, 1:7], y = training_set[, 8])
summary(loan_tree_model)
```

```{r}
# validation set
loan_tree_model_val <- C50::C5.0(x = validation_set[, 1:7], y = validation_set[, 8])
summary(loan_tree_model_val)
```

```{r}
# test set
loan_tree_model_test <- C50::C5.0(x = test_set[, 1:7], y = test_set[, 8])
summary(loan_tree_model_test)
```


The tree model simply classifies all loans as "can be fully paid" with an accuracy rate on training set of $1 - 14.4\% = 85.6\%$, which indeed achieves even higher accuracy rates on validation set and test set ($ 1 - 13.7\% = 86.3\%$ and $1 - 13.6\% = 86.4\%$) respectively. 

3. Experiment with different cost matrices to achieve a sensitivity (also known as *recall*) of approximately 25\%, 40\% and 50\% in your validation set. Also report the percentage of the loans $\frac {n11}{n11+n21}$ you would recommend to the bank for re-evaluation that were indeed charged off (also known as *precision*).

A sensitivity of 25\% (40\%, 50\%) means $\frac {\#\ of\ correctly\ predicted\ charged\ off\ loans}{\#\ of\ all\ actually\ charged\ off\ loans} = 25\%\ (40\%, 50\%)$, or $\#\ of\ correctly\ predicted\ charged\ off\ loans = \#\ of\ all\ actually\ charged\ off\ loans * 25\%\ (40\%, 50\%)$, or "there should be around 1 out of 4 (2.5, 2) actually charged off loans that is predicted correctly".

```{r}
# sensitivity = 25%, training
cost_mat1 <- matrix(c(NA, 1, 3.05, NA), 2, 2, byrow = TRUE)
rownames(cost_mat1) <- colnames(cost_mat1) <- c("Charged Off", "Fully Paid")
loan_tree_model1 <- C50::C5.0(x = training_set[, 1:7], y = training_set[, 8], costs = cost_mat1)
conf_mat1 <- confusionMatrix(data = predict(loan_tree_model1, training_set[, 1:7]), reference =  training_set[, 8])
conf_mat1$table
kable(conf_mat1$byClass, caption = "Important Ratios")
```

```{r}
# sensitivity = 25%, validation
pred1 <- predict(loan_tree_model1, validation_set[, 1:7])
confusionMatrix(data = pred1, reference = validation_set[, 8])
```


```{r, echo=TRUE}
# sensitivity = 40%, training
cost_mat2 <- matrix(c(NA, 1, 4, NA), 2, 2, byrow = TRUE)
rownames(cost_mat2) <- colnames(cost_mat2) <- c("Charged Off", "Fully Paid")
loan_tree_model2 <- C50::C5.0(x = training_set[, 1:7], y = training_set[, 8], costs = cost_mat2)
conf_mat2 <- confusionMatrix(data = predict(loan_tree_model2, training_set[, 1:7]), reference = training_set[, 8])
conf_mat2$table
kable(conf_mat2$byClass, caption = "Important Ratios")
```

```{r}
# sensitivity = 40%, validation
pred2 <- predict(loan_tree_model2, validation_set[, 1:7])
confusionMatrix(data = pred2, reference = validation_set[, 8])
```


```{r, echo=TRUE}
# sensitivity = 50%, training
cost_mat3 <- matrix(c(NA, 1, 4.9, NA), 2, 2, byrow = TRUE)
rownames(cost_mat3) <- colnames(cost_mat3) <- c("Charged Off", "Fully Paid")
loan_tree_model3 <- C50::C5.0(x = training_set[, 1:7], y = training_set[, 8], costs = cost_mat3)
conf_mat3 <- confusionMatrix(data = training_set[, 8], reference = predict(loan_tree_model3, training_set[, 1:7]))
conf_mat3$table
kable(conf_mat3$byClass, caption = "Important Ratios")
```

```{r}
# sensitivity = 50%, validation
pred3 <- predict(loan_tree_model3, validation_set[, 1:7])
confusionMatrix(data = pred3, reference = validation_set[, 8])
```


It is very important to identify the loans that might be defaulted in the future; on the other hand, classifying too many loans that are actually safe as dangerous would incur higher costs to deal with. We need to find a balance between high sensitivity as well as high specificity.

......


4. Pick a cost parameter matrix that you assess as the most appropriate for identifying loan applications that deserve further examination.

```{r}
sens_training <- vector(mode = 'double', length = 31)
sens_validation <- vector(mode = 'double', length = 31)
k_val <- vector(mode = 'double', length = 31)
for (k in 2:31) {
  cost_mat <- matrix(c(NA, 1, k, NA), 2, 2, byrow = TRUE)
  rownames(cost_mat) <- colnames(cost_mat) <- c("Charged Off", "Fully Paid")
  loan_tree_model <- C50::C5.0(x = training_set[, 1:7], y = training_set[, 8], costs = cost_mat)
  conf_mat_training <- confusionMatrix(data = training_set[, 8], reference = predict(loan_tree_model, training_set[, 1:7]))
  conf_mat_validation <- confusionMatrix(data = validation_set[, 8], reference = predict(loan_tree_model, validation_set[, 1:7]))
  sens_training[k] <- conf_mat_training$byClass['Sensitivity']
  sens_validation[k] <- conf_mat_validation$byClass['Sensitivity']
  k_val[k] <- k
}

df <- data.frame(k_val, sens_training, sens_validation)
df <- melt(df, id = "k_val")

ggplot(data = df, aes(x = k_val, y = value, col = variable)) +
  geom_point() +
  geom_line() +
  labs(title = 'Sensitivity vs Cost parameter', x = 'k: the cost parameter for misclassified charged-off loan (base: misclassified fully-paid loan as 1)', y = 'Sensitivity') +
  scale_fill_discrete(name = 'Dataset', labels = c('Training', 'Validation')) +
  theme(legend.position = "right")
```





5. Evaluate the performance of your cost parameter matrix on the test set.

```{r}
# sensitivity = 25%
pred1 <- predict(loan_tree_model1, test_set[, 1:7])
confusionMatrix(data = test_set[, 8], reference = pred1)
```

```{r}
# sensitivity = 40%
pred2 <- predict(loan_tree_model2, test_set[, 1:7])
confusionMatrix(data = test_set[, 8], reference = pred2)
```

```{r}
# sensitivity = 50%
pred3 <- predict(loan_tree_model3, test_set[, 1:7])
confusionMatrix(data = test_set[, 8], reference = pred3)
```



